---
title: "Bayes.Models"
output: html_document
date: "2025-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries

```{r}
library(tidyverse)
library(brms)
library(corrplot)
```

### Loading and correcting data with Brandie's code: https://github.com/BrandieQC/IPM_Practice/blob/main/IntBio_WL2_2023_Climate.Rmd

```{r}
growth <- read_csv("../Data/WL2-2023_Size_Combined.csv") %>% 
  select(survey_date, Genotype, parent.pop, height.cm, long.leaf.cm)

unique(growth$parent.pop) #all 23 pops!

surv <- read_csv("../Data/WL2_Mortality_2023.csv") %>% 
  select(block, BedLoc, Genotype, parent.pop=pop, mf:death.date) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% 
  mutate(Genotype= str_replace(Genotype, "Y08", "YO8")) %>% 
  mutate(Genotype= str_replace(Genotype, "Y04", "YO4")) 

unique(surv$parent.pop) #all 23 pops!
```

Change height column name to match common IPM terminology

```{r}
colnames(growth)[4] = "size"
```

GET RID OF DUPLICATES:

```{r}
growth.2 <- growth %>% 
  arrange(Genotype, survey_date, size) %>% 
  distinct(Genotype, survey_date, .keep_all=TRUE)
```

Merge with corrections 

```{r}
all.data <- left_join(growth.2, surv) %>%
  select(survey_date:parent.pop, mf:rep, block:BedLoc, size, long.leaf.cm, death.date) %>% 
  drop_na(block) #get rid of plants that were not planted and never got a block assigned. 

tail(all.data, 10)
unique(all.data$death.date)
```

Add in surv and sizenext columns

```{r}
all.data.timeprep <- all.data %>% 
  select(-long.leaf.cm) %>% 
  mutate(death.date=mdy(death.date)) %>% 
  group_by(Genotype) %>% 
  mutate(sizeNext = lead(size, order_by = survey_date), # next timepoints height
        # survey_date_char = as.character(survey_date), #grouping variable to compare to continuous survey date 
         date_next = lead(survey_date, order_by = survey_date), #needed for surv code and interval calc
         elapsed_days= date_next - survey_date, #interval length in days
         elapsed_weeks = as.integer(elapsed_days)/7, #interval length in weeks 
         surv=if_else(is.na(death.date), 1,
                      if_else(date_next>=death.date, 0, 1))) %>% #survival to the next size time point
  ungroup() %>%
  mutate(week = as.numeric(survey_date - ymd("2023-07-03")), #Weeks since pre-transplant size 
         pop.mf=str_c(parent.pop, "_", mf)) %>% #for specifying random effects separately  
  drop_na(size, sizeNext, surv) #for models
```

### Plotting

Plotting survival, growth/shrinkage/stasis as a function of our state variable, size.

```{r}
sizenext.plot=ggplot(all.data.timeprep, aes(x = size, y = sizeNext))+
  geom_point()+
  xlab("Size (t)")+
  ylab("Size to t+1")
sizenext.plot
```

```{r}
sizenext.pop.plot=ggplot(all.data.timeprep, aes(x = size, y = sizeNext))+
  geom_point()+
  xlab("Size (t)")+
  ylab("Size to t+1")+
  facet_wrap(~parent.pop)
sizenext.pop.plot
```

```{r}
all.data.timeprep %>%  #pretty skewed
  ggplot(aes(x=size)) +
  geom_histogram()
```

```{r}
all.data.timeprep %>%  #pretty skewed
  ggplot(aes(x=sizeNext)) +
  geom_histogram()
```

### Log Transformation

Log transform size and sizeNext

```{r}
all.data.timeprep = all.data.timeprep %>%
  mutate(size.log = log(size),
         sizeNext.log = log(sizeNext))
```

```{r}
all.data.timeprep %>%  #pretty skewed
  ggplot(aes(x=size.log)) +
  geom_histogram()
```

```{r}
all.data.timeprep %>%  #pretty skewed
  ggplot(aes(x=sizeNext.log)) +
  geom_histogram()
```

### Add in weekly climate data

Soil Water Content: Soil moisture

```{r}
swc_raw <- read_csv("../Data/WL2_2023_Bed_C_Soil_Moisture_Corrected.csv") %>% 
  mutate(Date_Time=mdy_hm(Date_Time))
head(swc_raw)

swc_port_avgs <- swc_raw %>% #avg across the probes (all in same bed)
  rowwise() %>% #for each row...
  mutate(SWC_port_avg = mean(c_across(-Date_Time))) %>%  #avg across all columns except Date_Time
  select(Date_Time, SWC_port_avg) %>%
  mutate(Date=as.Date(Date_Time)) 

swc_daily_summary <- swc_port_avgs %>%
  group_by(Date) %>%
  summarize(
    Daily_AvgSWC = mean(SWC_port_avg),
    Daily_MaxSWC = max(SWC_port_avg),
    Daily_MinSWC = min(SWC_port_avg))
head(swc_daily_summary)
summary(swc_daily_summary)
```

Soil Temperature

```{r}
soil_temp_raw <- read_csv("../Data/WL2_2022_2023_iButton_Data_Corrected.csv") %>% 
  mutate(Date_Time=mdy_hm(Date_Time))
head(soil_temp_raw)

unique(soil_temp_raw$Bed) #probes not spread out enough to assign half of the field to one probe and the other half to another...
#A_2 kind of captures the variance and the trends of the other probes 

soil_temp_bedA <- soil_temp_raw %>% 
  filter(Bed=="A_2") %>% 
  separate(Date_Time, into = c("Date", "Time"), sep = " ") %>% 
   mutate(Date=as.Date(Date)) %>% 
  select(Bed:Time, SoilTemp)

soil_temp_daily_summary <- soil_temp_bedA %>% 
  group_by(Date) %>%
  summarize(
    Daily_AvgTemp = mean(SoilTemp),
    Daily_MaxTemp = max(SoilTemp),
    Daily_MinTemp = min(SoilTemp),
    Diurnal_Range = Daily_MaxTemp-Daily_MinTemp)
head(soil_temp_daily_summary)
summary(soil_temp_daily_summary)
```

### Correlation between SWC and soil temp

```{r}
daily_temp_swc <- full_join(swc_daily_summary, soil_temp_daily_summary) %>% 
  drop_na() #drop NAs - note you lose some dates for both temp and swc 
summary(daily_temp_swc)

temp_swc_scaled <- daily_temp_swc %>% select(Daily_AvgSWC:Diurnal_Range) %>% scale() #normalize the data so they're all on the same scale
#head(temp_swc_scaled)
cor.norm_tempswc = cor(temp_swc_scaled) #test correlations among the traits
cor.sig_tempswc <- cor.mtest(temp_swc_scaled, method="pearson") #test significance of corrs
corrplot(cor.norm_tempswc, type = "upper",
         tl.srt = 45, p.mat = cor.sig_tempswc$p, 
         sig.level = 0.05, insig="blank") # a lot of significant correlations! --> keep separate 

cor.norm_tempswc
#use only daily AVG SWC b/c highly correlated to max and min swc (r > 0.96)
#avg and max temp highly correlated (r=0.93)
#max temp and diurnal range highly correlated (r=0.95) --> don't use max temp 
```

Brandie had SWC and temp. in separate models. I'm going to try Daily_AvgSWC and Daily_MinTemp b/c r = -0.38

### Merge with size/surv data 

```{r}
date_intervals <- all.data %>% 
  select(survey_date) %>% 
  distinct() %>% 
  mutate(prev_date = lag(survey_date, order_by = survey_date)) 
```

```{r}
swc_intervals <- swc_daily_summary %>% 
  mutate(Interval=if_else(Date>="2023-07-26" & Date<"2023-08-02", 1,
                                if_else(Date>="2023-08-02" & Date<"2023-08-16", 2,
                                if_else(Date>="2023-08-16" & Date<"2023-08-23", 3,
                                if_else(Date>="2023-08-23" & Date<"2023-08-30", 4,
                                if_else(Date>="2023-08-30" & Date<"2023-09-06", 5,
                                if_else(Date>="2023-09-06" & Date<"2023-09-13", 6,
                                if_else(Date>="2023-09-13" & Date<"2023-09-20", 7,
                                if_else(Date>="2023-09-20" & Date<"2023-09-27", 8,
                                if_else(Date>="2023-09-27" & Date<"2023-10-13", 9,
                                if_else(Date>="2023-10-13" & Date<"2023-10-20", 10, NA))))))))))) %>% 
  select(Date:Daily_AvgSWC, Interval) %>% 
  group_by(Interval) %>% 
  filter(!is.na(Interval)) %>% 
  summarise(meanSWC=mean(Daily_AvgSWC, na.rm=TRUE))
```

```{r}
soil_temp_intervals <- soil_temp_daily_summary %>% 
  mutate(Interval=if_else(Date>="2023-07-26" & Date<"2023-08-02", 1,
                                                    if_else(Date>="2023-08-02" & Date<"2023-08-16", 2,
                                                    if_else(Date>="2023-08-16" & Date<"2023-08-23", 3,
                                                    if_else(Date>="2023-08-23" & Date<"2023-08-30", 4,
                                                    if_else(Date>="2023-08-30" & Date<"2023-09-06", 5,
                                                    if_else(Date>="2023-09-06" & Date<"2023-09-13", 6,
                                                    if_else(Date>="2023-09-13" & Date<"2023-09-20", 7,
                                                    if_else(Date>="2023-09-20" & Date<"2023-09-27", 8,
                                                    if_else(Date>="2023-09-27" & Date<"2023-10-13", 9,
                                                    if_else(Date>="2023-10-13" & Date<"2023-10-20", 10, NA))))))))))) %>% 
  group_by(Interval) %>% 
  filter(!is.na(Interval)) %>% 
  summarise(meanTemp=mean(Daily_AvgTemp, na.rm=TRUE), maxTemp=max(Daily_MaxTemp, na.rm = TRUE), 
            minTemp=min(Daily_MinTemp, na.rm = TRUE), meanDiurnal_Range=mean(Diurnal_Range, na.rm = TRUE))
```

```{r}
all.data.clim.for.surv <- all.data.timeprep %>% 
  mutate(Interval=if_else(survey_date=="2023-07-26", 1,
                              if_else(survey_date=="2023-08-02", 2,
                              if_else(survey_date=="2023-08-16", 3,
                              if_else(survey_date=="2023-08-23", 4,
                              if_else(survey_date=="2023-08-30", 5,
                              if_else(survey_date=="2023-09-06", 6,
                              if_else(survey_date=="2023-09-13", 7,
                              if_else(survey_date=="2023-09-20", 8,
                              if_else(survey_date=="2023-09-27", 9,
                              if_else(survey_date=="2023-10-13", 10, NA))))))))))) %>% 
  left_join(soil_temp_intervals) %>% 
  left_join(swc_intervals) %>% 
  filter(!is.na(Interval))
head(all.data.clim.for.surv)
summary(all.data.clim.for.surv)
names(all.data.clim.for.surv)
```

```{r}
all.data.clim.for.growth <- all.data.timeprep %>% 
  mutate(Interval=if_else(survey_date=="2023-07-26", 1,
                              if_else(survey_date=="2023-08-02", 2,
                              if_else(survey_date=="2023-08-16", 3,
                              if_else(survey_date=="2023-08-23", 4,
                              if_else(survey_date=="2023-08-30", 5,
                              if_else(survey_date=="2023-09-06", 6,
                              if_else(survey_date=="2023-09-13", 7,
                              if_else(survey_date=="2023-09-20", 8,
                              if_else(survey_date=="2023-09-27", 9,
                              if_else(survey_date=="2023-10-13", 10, NA))))))))))) %>% 
  left_join(soil_temp_intervals) %>% 
  left_join(swc_intervals) %>% 
  filter(!is.na(Interval))
head(all.data.clim.for.growth)
summary(all.data.clim.for.growth)
names(all.data.clim.for.growth)
```

### Survival Models

Instead of weeks or aspect of time in the model, I am going to use an observation random effect (OBRE).
We need OBRE b/c measure same individuals multiple times over time

```{r}
priors <- c(prior(normal(0, 10), class = b))

linear.mod = brm(surv ~ 1 + size.log + (1|block) + (1|parent.pop/mf) + (1|Genotype),
                 family = bernoulli(),
                 data = all.data.clim.for.surv)
linear.clim.mod = brm(surv ~ 1 + size.log + minTemp + meanSWC + (1|block) + (1|parent.pop/mf) 
                      + (1|Genotype),
                 family = bernoulli(),
                 data = all.data.clim.for.surv)
quad.mod = brm(surv ~ 1 + size.log + I(size.log^2) + (1|block) + (1|parent.pop/mf) 
               + (1|Genotype),
                family = bernoulli(),
                 data = all.data.clim.for.surv)
quad.clim.mod = brm(surv ~ 1 + size.log + I(size.log^2) + minTemp + meanSWC + (1|block) 
                    + (1|parent.pop/mf) + (1|Genotype),
                 family = bernoulli(),
                 data = all.data.clim.for.surv)
cubic.mod = brm(surv ~ 1 + size.log + I(size.log^2) + I(size.log^3) + (1|block) 
                + (1|parent.pop/mf) + (1|Genotype),
                family = bernoulli(),
                 data = all.data.clim.for.surv)
cubic.clim.mod = brm(surv ~ 1 + size.log + I(size.log^2) + I(size.log^3) + minTemp + meanSWC 
                    + (1|block) + (1|parent.pop/mf) + (1|Genotype),
                 family = bernoulli(),
                 data = all.data.clim.for.surv)

```

Model Comparison

```{r}
linear.mod = add_criterion(linear.mod, c("loo", "waic"))
linear.clim.mod = add_criterion(linear.clim.mod, c("loo", "waic"))
quad.mod = add_criterion(quad.mod, c("loo", "waic"))
quad.clim.mod = add_criterion(quad.clim.mod, c("loo", "waic"))
cubic.mod = add_criterion(cubic.mod, c("loo", "waic"))
cubic.clim.mod = add_criterion(cubic.clim.mod, c("loo", "waic"))
```

```{r}
summary(linear.mod)
summary(linear.clim.mod)
summary(quad.mod)
summary(quad.clim.mod)
summary(cubic.mod)
summary(cubic.clim.mod)
```

Missing interval 10 for soil temperature

```{r}
loo_compare(linear.mod,quad.mod,cubic.mod, criterion = "loo")
```

```{r}
loo_compare(linear.clim.mod,quad.clim.mod,cubic.clim.mod, criterion = "loo")
```


```{r}
loo_compare(linear.mod,linear.clim.mod,quad.mod,quad.clim.mod,cubic.mod,cubic.clim.mod, criterion = "waic")
```

```{r}
loo_compare(linear.mod,linear.clim.mod,quad.mod,quad.clim.mod,cubic.mod,cubic.clim.mod, criterion = "loo")
```










