---
title: "Real.Data"
output: html_document
date: "2025-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(magrittr)
```

Read in the growth and survival data

```{r}
growth = read_csv("../Data/WL2-2023_Size_Combined.csv")
surv = read_csv("../Data/WL2_Mortality_2023.csv")
```

Columns we need in the final data frame for analysis: size, sizeNext, surv (0,1), fec.seed (number of seeds), fec.flower (flowering status,0,1). If suv = 0, then sizeNext is NA.

```{r}
head(growth)
head(surv)
```

Correcting duplication in growth

```{r}
growth.2 = growth %>%
  arrange(Genotype, survey_date, height.cm) %>%
  distinct(Genotype, survey_date, .keep_all = TRUE)
```

Change column names to match common IPM terminology

```{r}
colnames(growth.2)[8] = "size"
```

Wrangle growth data using Julin's code

```{r}
growth.3 <- growth.2 %>% 
  arrange(Genotype, survey_date) %>%
  select(-survey.notes, -long.leaf.cm) %>%
  group_by(Genotype) %>%
  mutate(sizeNext = lead(size),
         elapsed_days= lead(survey_date) - survey_date, # works b/c of read_csv, reads in survey_date as Date
         elapsed_weeks = as.integer(elapsed_days)/7,
         survey_date_char = as.character(survey_date)) %>%
  ungroup() %>%
  drop_na(size, sizeNext)

growth.3 
```

Merge survival and growth data. Remove block from growth first.

```{r}
growth.4 = growth.3 %>%
  select(Genotype,parent.pop,mf,rep,size,sizeNext,elapsed_weeks,survey_date_char)%>%
  rename(pop = parent.pop)
```

```{r}
all.data = merge(growth.4,surv)
```


### Plotting

Plotting survival, growth/shrinkage/stasis as a function of our state variable, size.

```{r}
sizenext.plot=ggplot(growth.4, aes(x = size, y = sizeNext))+
  geom_point()+
  xlab("Size (t)")+
  ylab("Size to t+1")
sizenext.plot
```

```{r}
sizenext.pop.plot=ggplot(growth.4, aes(x = size, y = sizeNext))+
  geom_point()+
  xlab("Size (t)")+
  ylab("Size to t+1")+
  facet_wrap(~pop)
sizenext.pop.plot
```

### Growth Kernel

Never finished running
```{r}

growth.models <- tibble(name=c("0_intercept", "1_linear", "2_linear_weeks","3_linear_date","4_quadratic", "5_quadratic_weeks",
                               "6_quadratic_date","7_cubic","8_cubic_weeks","9_cubic_date"),
                      f = c("sizeNext ~ 1 + (1|block) + (1|pop)", 
                            "sizeNext ~ size + (1|block) + (1+size|pop)", 
                            "sizeNext ~ size*elapsed_weeks + (1|block) + (1+size*elapsed_weeks|pop)", 
                            "sizeNext ~ size*survey_date_char + (1|block) + (1+size*survey_date_char|pop)", 
                             "sizeNext ~ size + I(size^2) + (1|block) + (1+size + I(size^2)|pop)", 
                            "sizeNext ~ size*elapsed_weeks + I(size^2)*elapsed_weeks + (1|block) + (1+size*elapsed_weeks + I(size^2)*elapsed_weeks|pop)", 
                            "sizeNext ~ size*survey_date_char + I(size^2)*survey_date_char + (1|block) + (1+size*survey_date_char + I(size^2)*survey_date_char|pop)",
                            "sizeNext ~ size + I(size^2) + I(size^3) + (1|block) + (1|pop)",
                            "sizeNext ~ size*elapsed_weeks + I(size^2)*elapsed_weeks + I(size^3)*elapsed_weeks + (1|block) + (1+size*elapsed_weeks + I(size^2)*elapsed_weeks + I(size^3)*elapsed_weeks|pop)",
                            "sizeNext ~ size*survey_date_char + I(size^2)*survey_date_char + I(size^3)*survey_date_char + (1|block) + (1+size*survey_date_char + I(size^2)*survey_date_char + I(size^3)*survey_date_char|pop)"))

growth.models.2 <- growth.models %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), data = all.data)),
         predict = map(lmer, ~ cbind(all.data, preds=predict(.))),
         glance = map(lmer, broom.mixed::glance))

```

```{r}

growth.models <- tibble(name=c("0_intercept", "1_linear", "2_linear_weeks","4_quadratic", "5_quadratic_weeks",
                               "7_cubic","8_cubic_weeks"),
                      f = c("sizeNext ~ 1 + (1|block) + (1|pop)", 
                            "sizeNext ~ size + (1|block) + (1+size|pop)", 
                            "sizeNext ~ size*elapsed_weeks + (1|block) + (1+size*elapsed_weeks|pop)", 
                             "sizeNext ~ size + I(size^2) + (1|block) + (1+size + I(size^2)|pop)", 
                            "sizeNext ~ size*elapsed_weeks + I(size^2)*elapsed_weeks + (1|block) + (1+size*elapsed_weeks + I(size^2)*elapsed_weeks|pop)", 
                            "sizeNext ~ size + I(size^2) + I(size^3) + (1|block) + (1|pop)",
                            "sizeNext ~ size*elapsed_weeks + I(size^2)*elapsed_weeks + I(size^3)*elapsed_weeks + (1|block) + (1+size*elapsed_weeks + I(size^2)*elapsed_weeks + I(size^3)*elapsed_weeks|pop)"))

growth.models.2 <- growth.models %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), data = all.data)),
         predict = map(lmer, ~ cbind(all.data, preds=predict(.))),
         glance = map(lmer, broom.mixed::glance))

```



Linear best fit according to AIC

```{r}
growth.models %>% unnest(glance)
```

```{r}
growth.model.final = growth.models %>% filter(name == "1_linear") %>% pull(lmer) %>% magrittr::extract2(1)
summary(growth.model.final)
```

Plot the predictions for the linear model.

```{r}
growth.predictions <- growth.models %>% dplyr::select(name, predict) %>% unnest(predict)

ggplot(growth.predictions, aes(x = size, y = sizeNext)) +
  geom_point()+
  geom_smooth(aes(y = preds, color = name))+
  scale_color_viridis_d(end = 0.9)
```


### Survival Kernel

```{r}

#newdata <- expand_grid(
  #parent.pop = unique(growth.6$parent.pop),
  #pop.mf.2 = unique(growth.6$pop.mf.2),
  #block = unique(growth.6$block),
  #size = seq(min(growth.6$size, na.rm = TRUE), max(growth.6$size, na.rm = TRUE), length.out = 100))

surv.models <- tibble(name=c("0_intercept", "1_linear", "2_quadratic", "3_cubic"),
                      f = c("surv ~ 1 + (1|block) + (1|parent.pop/pop.mf)", 
                            "surv ~ size + (1|block) + (1|parent.pop/pop.mf)", 
                            "surv ~ size + I(size^2) + (1|block) + (1|parent.pop/pop.mf)", 
                            "surv ~ size + I(size^2) + I(size^3) + (1|block) + (1|parent.pop/pop.mf)"))

surv.models <- surv.models %>%
  mutate(glmer = map(f, ~ glmer(as.formula(.), data = growth.7, family = "binomial")),
         predict = map(glmer, ~ cbind(growth.7, preds=predict(., type = "response"))),
         glance = map(glmer, tidy),
         AIC = map(glmer, AIC))
```

Quadratic best fit according to AIC

```{r}
surv.models %>% unnest(AIC)
```

```{r}
survival.model.final = surv.models %>% filter(name == "3_cubic") %>% pull(glmer) %>% magrittr::extract2(1)
summary(survival.model.final)
```


Plot the predictions for the cubic model.

```{r}
surv.predictions <- surv.models %>% dplyr::select(name, predict) %>% unnest(predict)

ggplot(surv.predictions, aes(x = size, y = surv)) +
  geom_point()+
  geom_smooth(aes(y = preds, color = name))+
  scale_color_viridis_d(end = 0.9)
```



### Creating the P matrix

Create dataframe for survival and growth parameters

```{r}
params=data.frame(
  surv.int=NA, # Intercept from logistic regression of survival
  surv.slope1=NA, # Slope from logistic regression of survival
  surv.slope2=NA, # Quadratic slope from logistic regression of survival
  surv.slope3=NA, # Cubic slope from logistic regression of survival
  growth.int=NA, # Intercept from linear regression of growth
  growth.slope=NA, # Slope from linear regression of growth
  growth.sd=NA # Residual sd from the linear regression of growth
)
```

Get survival coefficients for cubic model

```{r}
surv.coefs <- surv.models %>% 
  filter(name == "3_cubic") %>%
  mutate(coefs = map(glmer, tidy)) %>% 
  dplyr::select(name, f, coefs) %>% 
  unnest(coefs) 

params$surv.int <- surv.coefs %>% filter(term == "(Intercept)") %>% pull(estimate)
params$surv.slope1 <- surv.coefs %>% filter(term == "size") %>% pull(estimate)
params$surv.slope2 <- surv.coefs %>% filter(term == "I(size^2)") %>% pull(estimate)
params$surv.slope3 <- surv.coefs %>% filter(term == "I(size^3)") %>% pull(estimate)
```

Get growth coefficients from linear model

```{r}
growth.coefs <- growth.models %>% 
  filter(name == "1_linear") %>%
  mutate(coefs = map(lmer, tidy)) %>% 
  dplyr::select(name, f, coefs) %>% 
  unnest(coefs)

params$growth.int <- growth.coefs %>% filter(term=="(Intercept)") %>% pull(estimate)
params$growth.slope <- growth.coefs %>% filter(term=="size") %>% pull(estimate)
params$growth.sd = sd(resid(growth.models$lmer[[2]])) # observed standard deviation
```

Set up the functions

```{r}
# survival probability function, this is a logistic function, provides survival probability based on size.
s.x=function(x,params) {
  u=exp(params$surv.int + params$surv.slope1*x + params$surv.slope2*x^2 + params$surv.slope3*x^3)
  return(u/(1+u))  # this is back transforming from logit
}

# growth function, this is a normal distribution, returns a probability distribution of new sizes at t+1 (xp) at a given size x.  
g.yx=function(xp,x,params) {
  dnorm(xp,mean=params$growth.int+params$growth.slope*x,sd=params$growth.sd)
}
```

Make the kernel and define the parameters for integration

```{r}
minSize<-min(growth.7$size,na.rm=T)
maxSize<-max(growth.7$size,na.rm=T)
nBigMatrix = 80

b <- minSize + c(0:nBigMatrix) * (maxSize - minSize)/nBigMatrix # boundary points
y <- 0.5 * (b[1:nBigMatrix] + b[2:(nBigMatrix + 1)]) # mesh points
h <- y[2] - y[1] # step size
```

Make the matrix

```{r}
G = h*outer(y,y,g.yx,params = params) # growth matrix
# OK for each value of y,y evaluate the growth function g.yx using params
# If I understand this correctly, the rows are x(t) and the columns are x(t+1)

S = s.x(y, params = params) # survival
P = G # placeholder, redefine P on the next line
for(i in 1:nBigMatrix) P[,i] = G[,i]*S[i] # growth/survival matrix
```

Correct the matrix using the Merow et al "constant" correction code

```{r}
Pc = P # making a corrected matrix from already generated matrix
nvals <- colSums(Pc, na.rm = TRUE)
loc0 <- which(nvals == 0, arr.ind = TRUE)
if (length(loc0) > 0) {
      print("warnings - columns that sum to 0 or that have NAs - assuming survival is along the diagonal; plot your Pmatrix to check it")
      Pc[, loc0] <- 0
      Pc[cbind(loc0, loc0)] <- s.x(size = y[loc0], params = params)
}
nvals <- colSums(Pc, na.rm = TRUE)
Pc <- t((t(Pc)/nvals) * s.x(y, params=params))
```

Plot the corrected matrix

```{r}
Pc %>% as_tibble() %>%
  set_colnames(y) %>% 
  mutate(size.t1=y) %>%
  pivot_longer(-size.t1, names_to = "size.t", names_transform = as.numeric) %>%
  ggplot(aes(x=size.t, y = size.t1)) +
  geom_raster(aes(fill = value)) +
  geom_contour(aes(z = value),lwd=.25) +
  geom_abline(intercept=0, slope = 1, color="gray90", lty=5) +
  scale_fill_viridis_c(option = "plasma") +
  labs(x = "Size (t)", y = "Size (t + 1)", title = "P Matrix (corrected): Size and Growth") +
  coord_equal() +
  theme_bw()
```
